---
title: "FPLC Unicorn File Analysis"
Author: "Sarah Lucas"
---
# Steps to get input data file

Step 1. Download PyCorn. Most computers already have python (at least 2.7, but probably 2.7/3.x) pre-installed so copy and past the following code into a Terminal:

pip install pycorn

Step 2. Export each Unicorn file as a .zip file and transfer them to your computer. Save all of them in a single folder. Enter that folder in your terminal. There are options for exporting files directly to .csv files, but I haven't been able to make it work yet. the pycorn-bin.py does havea csv flag instead of xlsx, but I haven't been able to make that work. So, Copy and past the following into your terminal with *zip ("all .zip files") as the 'input.res' option:

##Extract data to xlsx-file:
pycorn-bin.py -e xlsx input.res <-- DUMMY EXAMPLE
pycorn-bin.py -e xlsx *zip <-- COPY AND PASTE

Then, you can batch convert the xlsx files 

##download the python package xlsx2csv
pip install xlsx2csv

## make a script named batch_xlsx2csv.sh containing the text below and save it somewhere in your PATH (don't forget to de-comment the lines).
for i in *.xlsx;
 do
  filename=$(basename "$i" .xlsx);
  outext=".csv"
  xlsx2csv $i $filename$outext
done

## make the script executable
chmod +x batch_xlsx2csv.sh to 

## Go back to the folder with your new xlsx files and run the line
batch_xlsx2csv.sh

Now you should have a folder filled with the original .zip files, .xlsx files, and the new .csv files.

Step 3. Load the libararies below.
```{r}
library(DescTools)
library(tidyverse)
```
Aesthetics
```{r}

cbPal_MMM <- c("black","pink","#E69F00","#56B4E9","#009E73","#F0E442", "#0072B2","#75437C","#B52F25") 

theme_transparent <- theme(panel.background = element_rect(fill = "transparent"),
                           plot.background = element_rect(fill = "transparent", 
                                                        color = NA),
                           panel.grid.major = element_line(color = "transparent"),
                           panel.grid.minor = element_line(color = "transparent"),
                           legend.background = element_rect(fill = "transparent"), 
                           legend.box.background = element_rect(fill = "transparent", 
                                                                color = "transparent"))
```

Step 4a. Function for importing .csv files
```{r}
read_chr_data <- function (x) {
  read_csv(x, col_types = 
             cols(
               UV = col_double(),
               X18 = col_double()
             )) %>%
  data.frame() %>%
  select(ElutionVolume = UV, Absorbance280 = X18) %>%
 # filter(row_number() > 46, row_number() < 200)
  #filter(row_number() != 1) %>%
  slice(49:39797)
#  x$Absorbance280 <- as.numeric()
}
```
Step 4b. Import csv files. Customize to file location and data object naming.
```{r}
chr_data_MMM <- read_chr_data("~/Documents/MICaB/Hunter_Lab/Projects/inProgress/CRS_MMM_Enrichments/FPLC/FPLC_MMM_SUPERS/September2019Supernatants/MMM_blank.csv")
chr_data_164 <- read_chr_data("~/Documents/MICaB/Hunter_Lab/Projects/inProgress/CRS_MMM_Enrichments/FPLC/FPLC_MMM_SUPERS/September2019Supernatants/Super164_blank.csv")
chr_data_222 <- read_chr_data("~/Documents/MICaB/Hunter_Lab/Projects/inProgress/CRS_MMM_Enrichments/FPLC/FPLC_MMM_SUPERS/September2019Supernatants/Super222_blank.csv")
chr_data_230 <- read_chr_data("~/Documents/MICaB/Hunter_Lab/Projects/inProgress/CRS_MMM_Enrichments/FPLC/FPLC_MMM_SUPERS/September2019Supernatants/Super230_blank.csv")
chr_data_234 <- read_chr_data("~/Documents/MICaB/Hunter_Lab/Projects/inProgress/CRS_MMM_Enrichments/FPLC/FPLC_MMM_SUPERS/September2019Supernatants/Super234_blank.csv")
chr_data_235 <- read_chr_data("~/Documents/MICaB/Hunter_Lab/Projects/inProgress/CRS_MMM_Enrichments/FPLC/FPLC_MMM_SUPERS/September2019Supernatants/Super235_blank.csv")
chr_data_243 <- read_chr_data("~/Documents/MICaB/Hunter_Lab/Projects/inProgress/CRS_MMM_Enrichments/FPLC/FPLC_MMM_SUPERS/September2019Supernatants/Super243_blank.csv")
chr_data_260 <- read_chr_data("~/Documents/MICaB/Hunter_Lab/Projects/inProgress/CRS_MMM_Enrichments/FPLC/FPLC_MMM_SUPERS/September2019Supernatants/Super260_blank.csv")
chr_data_262 <- read_chr_data("~/Documents/MICaB/Hunter_Lab/Projects/inProgress/CRS_MMM_Enrichments/FPLC/FPLC_MMM_SUPERS/September2019Supernatants/Super262_blank.csv")
```

```{r}
#chr_data_MMM$ElutionVolume <- as.double(chr_data_MMM$ElutionVolume)
```


Step 5. Customizable. Add/delete data objects from the plot. Change colors, aesthetics, etc.
```{r}
#fig.height=5, fig.width=8
chromatogram <- ggplot(chr_data_MMM, aes(x=ElutionVolume, y = Absorbance280)) +
  geom_line(size = 1) +
  geom_step(data = chr_data_164, color = "pink") +
  geom_step(data = chr_data_222, color = "#E69F00") +
  geom_step(data = chr_data_230, color = "#56B4E9") +
  geom_step(data = chr_data_234, color = "#009E73") +
  geom_step(data = chr_data_235, color = "#F0E442") +
  geom_step(data = chr_data_243, color = "#0072B2") +
  geom_step(data = chr_data_260, color = "#75437C") +
  geom_step(data = chr_data_262, color = "#B52F25") +
#  geom_vline(xintercept = 6.8, alpha = 0.5, linetype = 2) +
#  geom_vline(xintercept = 8.8, alpha = 0.5, linetype = 2) +
  scale_x_continuous(name = "Elution Volume (mL)", 
#                     breaks = c(0,5,10,15,20,25)
                     ) +
  ylab("Absorbance (280nm)") +
  theme_pubr()
chromatogram
```

```{r}
chr_data_MMM_first_peak <- chr_data_MMM %>% filter(between(ElutionVolume, 7, 10))
chr_data_164_first_peak <- chr_data_164 %>% filter(between(ElutionVolume, 7, 10))
chr_data_222_first_peak <- chr_data_222 %>% filter(between(ElutionVolume, 7, 10))
chr_data_230_first_peak <- chr_data_230 %>% filter(between(ElutionVolume, 7, 10))
chr_data_234_first_peak <- chr_data_234 %>% filter(between(ElutionVolume, 7, 10))
chr_data_235_first_peak <- chr_data_235 %>% filter(between(ElutionVolume, 7, 10))
chr_data_243_first_peak <- chr_data_243 %>% filter(between(ElutionVolume, 7, 10))
chr_data_260_first_peak <- chr_data_260 %>% filter(between(ElutionVolume, 7, 10))
chr_data_262_first_peak <- chr_data_262 %>% filter(between(ElutionVolume, 7, 10))
```



```{r fig.height=3, fig.width=3}
chromatogram_firstpeak <- ggplot(chr_data_MMM_first_peak, aes(x=as.numeric(ElutionVolume), y = as.numeric(Absorbance280))) +
  geom_line() +
  geom_step(data = chr_data_164_first_peak, color = "pink") +
  geom_step(data = chr_data_222_first_peak, color = "#E69F00") +
  geom_step(data = chr_data_230_first_peak, color = "#56B4E9") +
  geom_step(data = chr_data_234_first_peak, color = "#009E73") +
  geom_step(data = chr_data_235_first_peak, color = "#F0E442") +
  geom_step(data = chr_data_243_first_peak, color = "#0072B2") +
  geom_step(data = chr_data_260_first_peak, color = "#75437C") +
  geom_step(data = chr_data_262_first_peak, color = "#B52F25") +
  geom_vline(xintercept = 7.2, alpha = 0.5, linetype = 2) +
  geom_vline(xintercept = 8.2, alpha = 0.5, linetype = 2) +
  scale_x_continuous(name = "Elution Volume (mL)", 
#                     breaks = c(0,5,10,15,20,25)
                     ) +
  ylab("Absorbance (280nm)") +
  theme_pubr(base_size = 10)
  theme(
    axis.ticks = element_line(color = "black"),
    panel.border = element_rect(fill = "transparent", color = "black")) +
    theme_transparent

chromatogram_firstpeak
```
```{r}
#ggsave(plot = chromatogram_firstpeak, file = "~/Documents/MICaB/Hunter_Lab/Projects/inProgress/CRS_MMM_Enrichments/FPLC/FPLC_MMM_SUPERS/September2019Supernatants/firstpeak.pdf", device = "pdf", width = 2.5, height = 2.5, useDingbats = FALSE, bg = "transparent")
#ggsave(plot = chromatogram_firstpeak, file = "~/Documents/MICaB/Hunter_Lab/Projects/inProgress/CRS_MMM_Enrichments/FPLC/FPLC_MMM_SUPERS/September2019Supernatants/firstpeak.png", device = "png", width = 2.5, height = 2.5, bg = "transparent")
```

Step 6. Calculate area under the curve (AUC)
Using this method: https://rdrr.io/cran/DescTools/man/AUC.html
Area under the curve calculation:
Create a function to apply to each dataset
```{r}
find_AUC <- function(data, upper = 7.2, lower = 8.2){
  data$ElutionVolume <- as.numeric(data$ElutionVolume)
  data$Absorbance280 <- as.numeric(data$Absorbance280)
  
  data_peak1 <- data %>%
  filter(ElutionVolume > upper) %>%
  filter(ElutionVolume < lower)

  
  x = data_peak1$ElutionVolume
  y = data_peak1$Absorbance280
  
  AUC(x = x, y = y, method = "spline", na.rm = FALSE)
}
```
Apply function to datasets. 
```{r}
AUC_MMM <- find_AUC(data = chr_data_MMM)
AUC_164 <- find_AUC(data = chr_data_164)
AUC_222 <- find_AUC(data = chr_data_222)
AUC_230 <- find_AUC(data = chr_data_230)
AUC_234 <- find_AUC(data = chr_data_234)
AUC_235 <- find_AUC(data = chr_data_235)
AUC_243 <- find_AUC(data = chr_data_243)
AUC_260 <- find_AUC(data = chr_data_260)
AUC_262 <- find_AUC(data = chr_data_262)

AUC <- data.frame(sample = c("MMM","164","222","230","234","235","243","260","262"), AUC = c(AUC_MMM, AUC_164, AUC_222, AUC_230, AUC_234, AUC_235, AUC_243, AUC_260, AUC_262))
```

Graph AUC as a bar graph
```{r, fig.height = 3, fig.width = 3}
cbPal_MMM <- c("black","pink","#E69F00","#56B4E9","#009E73","#F0E442", "#0072B2","#75437C","#B52F25") 

AUC_bargraph <- ggplot(AUC, aes(x = reorder(sample, -AUC), y = as.double(AUC), fill = sample, color = "black")) + geom_bar(stat = "identity", alpha = 1) +
  theme_pubr(base_size = 10) +
  ylab("AUC") +
  scale_x_discrete(breaks = c("MMM","164","222","230","234","235","243","260","262"),
                   labels = c("MMM", "A","B","C","D","E","F","G","H")) +
  scale_fill_manual(values = cbPal_MMM,
                    limits = c("MMM","164","222","230","234","235","243","260","262")) +
  scale_color_manual(values = "black") +
  theme_pubr() +
  theme( 
    legend.position = "",
    legend.title = element_blank(),
    axis.title.x = element_blank()
  ) + theme_transparent
AUC_bargraph
```
```{r}
#ggsave(plot = AUC_bargraph, file = "~/Documents/MICaB/Hunter_Lab/Projects/inProgress/CRS_MMM_Enrichments/FPLC/FPLC_MMM_SUPERS/September2019Supernatants/firstpeakAUC.pdf", device = "pdf", width = 3, height = 3, bg = "transparent")
#ggsave(plot = AUC_bargraph, file = "~/Documents/MICaB/Hunter_Lab/Projects/inProgress/CRS_MMM_Enrichments/FPLC/FPLC_MMM_SUPERS/September2019Supernatants/firstpeakAUC.jpg", device = "jpg", width = 3, height = 3, bg = "transparent")
```
