---
title: "R Notebook"
---
# Setup
```{r}
library(tidyverse)
library(ggpubr)
library(phyloseq)
library(naniar)
```

# Load in data
## Sample Data from filtered phyloseq objects
The clinical data is sourced from the filtered dataset originating in the Preprocessing.Rmd notebook.
```{r}
filteredData <- readRDS("../data/PHYLOSEQ_DATA/phyloseq_filtered_MAR2020.rds")
clinicalData <- data.frame(sample_data(filteredData))
clinicalData
# Make it easier to find data from CRS/non-CRS samples
nonCRSData <- filter(clinicalData, DIAG_CRS == "Healthy")
CRSData <- filter(clinicalData, !DIAG_CRS == "Healthy")
```
## Sample Order from heatmap clustering in Abundance_Analysis_TopTaxa.Rmd
Importing taxa heatmap to extract rownames for ordering samples in plots
```{r}
heatmap.CRS.df <- readRDS("../data/heatmap_CRS_df.rds")
CRSsampleOrder <- colnames(heatmap.CRS.df)
```
```{r}
heatmap.NonCRS.df <- readRDS("../data/heatmap_NonCRS_df.rds")
NonCRSsampleOrder <- colnames(heatmap.NonCRS.df)
```


# Plotting summary clinical data
## Snot20 total
```{r fig.height=4, fig.width=3}
snot20 <- clinicalData %>%
  filter(SNOT20_COMPLETE == "Complete") %>%
  select(subject_id, DIAG_CRS, SNOT20_TOTAL) %>%
  filter(!is.na(SNOT20_TOTAL))

p <- ggboxplot(snot20, x = "DIAG_CRS", y = "SNOT20_TOTAL",
          color = "DIAG_CRS", palette = "jco",
          add = "jitter") + theme(legend.position = "") + 
  labs(x="", y="SNOT20 Total")
#  Add p-value
p + stat_compare_means(method = "t.test", label.y = 5)
```

## Sinus Factors
### CRS dataset
```{r}
sinusFactorsCRS <- clinicalData %>%
  select(SAMPLE_NAME, DIAG_CRS, contains("SINUS_FACTORS")) %>%
  select(-contains(c("FESS", "OTHER"))) %>%
  filter(!is.na(SINUS_FACTORS_COMPLETE)) %>%
  select(-SINUS_FACTORS_COMPLETE, -SINUS_FACTORS_POLYPOIDCHANGE) %>%
  dplyr::rename(Smoker = SINUS_FACTORS_SMOKER,
                Polyps = SINUS_FACTORS_POLYPS, 
                GERD = SINUS_FACTORS_GERD,
                Asthma = SINUS_FACTORS_ASTHMA,
                Allergies = SINUS_FACTORS_ALLERGIES) %>%
  pivot_longer(cols = c("Smoker","Polyps", "GERD", "Asthma", "Allergies"), names_to = "FACTOR", values_to = "boo") %>%
  mutate(BOOL = ifelse(boo == c(1), TRUE, FALSE)) %>%
  filter(DIAG_CRS == "CRS")
sinusFactorsCRS
```

```{r fig.width=8, fig.height=3}
#1.1
sinusfactorsTableCRS <- ggplot(sinusFactorsCRS, aes(x=SAMPLE_NAME, y=FACTOR)) + 
  geom_tile(colour = "whitesmoke",size=.5, aes(fill = BOOL)) +
  #geom_point(aes(color = BOOL)) + # Make the squares dots instead
  #facet_grid(cols = vars(DIAG_CRS), scales = "free_x", space = "free") +
  scale_x_discrete(breaks = CRSsampleOrder, 
                   limits = CRSsampleOrder,
                   expand = c(0,0)) + #Make the boxes square
  scale_y_discrete(expand = c(0,0)) + #Make the boxes square
  coord_fixed(ratio=1) +  #Make the boxes square
  scale_fill_manual(values = c("white", "black")) +
  #scale_fill_manual(values = c("white", "white")) + # Make the squares dots instead
  #scale_color_manual(values = c("white", "black")) + # Make the squares dots instead
  theme_minimal(base_size = 10) + 
  theme(legend.position = "",
        axis.text.x = element_blank(),
        #axis.text.x = element_text(angle=90),
        axis.ticks = element_blank(),
        axis.title = element_blank())
sinusfactorsTableCRS
```
```{r}
ggsave(sinusfactorsTableCRS, filename = "../figures/sinusFactorsTableCRS.pdf",
       device = "pdf",
       height = 3,
       width = 8)
```

### Non-CRS dataset
```{r}
sinusFactorsNonCRS <- clinicalData %>%
  select(SAMPLE_NAME, DIAG_CRS, contains("SINUS_FACTORS")) %>%
  select(-contains(c("FESS", "OTHER"))) %>%
  filter(!is.na(SINUS_FACTORS_COMPLETE)) %>%
  select(-SINUS_FACTORS_COMPLETE, -SINUS_FACTORS_POLYPOIDCHANGE) %>%
  dplyr::rename(Smoker = SINUS_FACTORS_SMOKER,
                Polyps = SINUS_FACTORS_POLYPS, 
                GERD = SINUS_FACTORS_GERD,
                Asthma = SINUS_FACTORS_ASTHMA,
                Allergies = SINUS_FACTORS_ALLERGIES) %>%
  pivot_longer(cols = c("Smoker","Polyps", "GERD", "Asthma", "Allergies"), names_to = "FACTOR", values_to = "boo") %>%
  mutate(BOOL = ifelse(boo == 1, TRUE, FALSE)) %>%
  filter(DIAG_CRS == "Healthy")
sinusFactorsNonCRS
```

```{r fig.height=2}
#1.1
sinusfactorsTableNonCRS <- ggplot(sinusFactorsNonCRS, aes(x=SAMPLE_NAME, y=FACTOR)) + 
  geom_tile(colour = "whitesmoke",size=.5, aes(fill = BOOL)) +
  #geom_point(aes(color = BOOL)) + # Make the squares dots instead
  scale_x_discrete(breaks = NonCRSsampleOrder, 
                   limits = NonCRSsampleOrder,
                   expand = c(0,0)) + #Make the boxes square
  scale_y_discrete(expand = c(0,0)) + #Make the boxes square
  coord_fixed(ratio=1) +  #Make the boxes square
  scale_fill_manual(values = c("white", "black")) +
  #scale_fill_manual(values = c("white", "white")) + # Make the squares dots instead
  #scale_color_manual(values = c("white", "black")) + # Make the squares dots instead
  #theme_minimal(base_size = 10) + 
  theme(legend.position = "",
        axis.text.x = element_blank(),
        #axis.text.x = element_text(angle=90),
        axis.ticks = element_blank(),
        axis.title = element_blank())
sinusfactorsTableNonCRS
```

##SNOT20 Vars
### CRS dataset
```{r}
# Define variables to keep
snot20PlotKeep <- c("blow nose", "runny nose", "thick nasal discharge", "post-nasal discharge", "ear pain", "ear fullness", "facial pain & pressure")
# Make table
snot20QsCRS <- clinicalData %>%
  data.frame() %>%
  select(SAMPLE_NAME, subject_id, DIAG_CRS, blow_nose:embarrassed) %>%
  dplyr::rename("blow nose" = blow_nose,
                "difficulty falling asleep" = difficulty_falling_asleep,
                "ear fullness" = ear_fullness,
                "ear pain" = ear_pain,
                "facial pain & pressure" = facial_pain_pressure,
                "frustrated/restless/irritable" = frustrated_restless_irritable,
                "lack sleep" = lack_sleep,
                "post-nasal discharge" = post_nasal_discharge,
                "reduced concentration" = reduced_concentration,
                "runny nose" = runny_nose,
                "thick nasal discharge" = thick_nasal_discharge,
                "wake tired" = wake_tired,
                "wake up at night" = wake_up_at_night
                ) %>%
  pivot_longer(cols = "blow nose":"embarrassed", names_to = "FACTOR", values_to = "VALUE") %>%
  replace_with_na(replace = list(VALUE = 0)) %>%
  #filter(FACTOR %in% snot20PlotKeep) %>%
  filter(DIAG_CRS == "CRS")
```
```{r}
snot20TableCRS <- ggplot(snot20QsCRS, aes(x=SAMPLE_NAME, y=FACTOR)) + 
  geom_tile(colour = "whitesmoke", size=.5, aes(fill=DIAG_CRS)) +
  geom_point(aes(size = VALUE, color = VALUE)) + # Make the squares dots instead
  scale_x_discrete(breaks = CRSsampleOrder, 
                   limits = CRSsampleOrder,
                   expand = c(0,0)) + #Make the boxes square
  scale_y_discrete(expand = c(0,0)) + #Make the boxes square
  coord_fixed(ratio=1) +  #Make the boxes square
  scale_size(range = c(.5,2), breaks = c(1,2,3,4,5)) +
  #scale_colour_brewer(palette = "Greens") +
  scale_fill_manual(values = c("white", "white")) +
  scale_color_gradient(low = "grey", 
                       high = "red") + 
  theme_minimal(base_size = 10) +
  theme(legend.position = "",
        axis.text.x = element_text(angle=90),
        #axis.text.x = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_blank())
snot20TableCRS
```
```{r}
ggsave(snot20TableCRS, filename = "../figures/snot20TableCRS.",
       device = "pdf",
       height = 3,
       width = 8)
```

### Non-CRS dataset
```{r}
snot20QsNonCRS <- clinicalData %>%
  data.frame() %>%
  select(SAMPLE_NAME, subject_id, DIAG_CRS, blow_nose:embarrassed) %>%
  dplyr::rename("blow nose" = blow_nose,
                "difficulty falling asleep" = difficulty_falling_asleep,
                "ear fullness" = ear_fullness,
                "ear pain" = ear_pain,
                "facial pain & pressure" = facial_pain_pressure,
                "frustrated/restless/irritable" = frustrated_restless_irritable,
                "lack sleep" = lack_sleep,
                "post-nasal discharge" = post_nasal_discharge,
                "reduced concentration" = reduced_concentration,
                "runny nose" = runny_nose,
                "thick nasal discharge" = thick_nasal_discharge,
                "wake tired" = wake_tired,
                "wake up at night" = wake_up_at_night
                ) %>%
  pivot_longer(cols = "blow nose":"embarrassed", names_to = "FACTOR", values_to = "VALUE") %>%
  replace_with_na(replace = list(VALUE = 0)) %>%
#  filter(FACTOR %in% snot20PlotKeep) %>%
  filter(DIAG_CRS == "Healthy")
```

```{r}
snot20TableNonCRS <- ggplot(snot20QsNonCRS, aes(x=SAMPLE_NAME, y=FACTOR)) + 
  geom_tile(colour = "whitesmoke", size=.5, aes(fill=DIAG_CRS)) +
  geom_point(aes(size = VALUE, color = VALUE)) + # Make the squares dots instead
  scale_x_discrete(breaks = NonCRSsampleOrder, 
                   limits = NonCRSsampleOrder,
                   expand = c(0,0)) + #Make the boxes square
  scale_y_discrete(expand = c(0,0)) + #Make the boxes square
  coord_fixed(ratio=1) +  #Make the boxes square
  #facet_grid(cols = vars(DIAG_CRS), scales = "free_x", space = "free") +
  scale_size(range = c(.5,2), breaks = c(1,2,3,4,5)) +
  scale_fill_manual(values = c("white", "white")) +
  scale_color_gradient(low = "grey",  
                       high = "red") + 
  theme(legend.position = "",
        #axis.text.x = element_blank(),
        axis.text.x = element_text(angle=90),
        axis.ticks = element_blank(),
        axis.title = element_blank())
snot20TableNonCRS
```
```{r}
ggsave(snot20TableNonCRS, filename = "../figures/snot20TableNonCRS.",
       device = "pdf",
       height = 3,
       width = 8)
```


Output a table of summary data for the following variables: Group by CRS status
Sex, Age, Polyps, Medications: Antibiotics/Steroids (#) Culture (#) Previous FESS (#), SNOT-22 measurements, 